---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Before starting with the analysis, these have been the options set and the operations made:

```{r echo=TRUE}
library(tidyverse)
options(max.print = 50, scipen = 999)

unzip("./activity.zip")
activity <- read.csv("./activity.csv")
activity$date <- parse_date(activity$date, "%Y-%m-%d")

```


## What is mean total number of steps taken per day?

Following command calculates sum of steps taken per day:
```{r echo=TRUE}
steps_by_day <- activity %>%
  group_by(date) %>%
  summarize(
    total_steps = sum(steps, na.rm = TRUE)
  )
```

Histogram of steps taken per day:
```{r echo=TRUE, fig.height=4}
ggplot(steps_by_day, aes(total_steps))+
  geom_histogram()+
  theme_minimal()+
  labs(x="Steps Taken", y="Number Days", title="Number of Days with same Step Count")
```

```{r echo=TRUE}
mean_steps_by_date <- sum(activity$steps,na.rm = TRUE)/length(unique(activity$date))

median_steps_by_date <- median(steps_by_day$total_steps,na.rm=TRUE)
```

The mean is `r mean_steps_by_date` and the median is `r median_steps_by_date`

## What is the average daily activity pattern?

Temporal distribution of steps taken during a day
```{r echo=TRUE, fig.height=4}
steps_by_interval <- activity %>%
  group_by(interval)%>%
  summarize(
    avg_steps = mean(steps,na.rm = TRUE)
  )

ggplot(steps_by_interval,aes(interval,avg_steps))+
  geom_line()+
  theme_minimal()+
  labs(x="Time Interval",y="Average Steps","Average Steps per Time Interval")
```


```{r echo=TRUE}
max_interval <- arrange(steps_by_interval,desc(avg_steps))[1,1]
```

The 5-minute interval that cointains the maximum number of steps is `r max_interval`

## Imputing missing values


```{r echo=TRUE}
total_nas <- sum(is.na(activity$steps))
```

Total missing values were `r total_nas`

Missing values correction was made by using average steps per interval:
```{r echo=TRUE}
activity_revised <- activity
  
for(i in 1:length(activity_revised$steps)){
  if(is.na(activity_revised$steps[i]) == TRUE) {
    activity_revised$steps[i] = filter(steps_by_interval, interval == activity_revised$interval[i])[[1,2]]
  } 
}
```

Distribution of the new data:
```{r echo=TRUE, fig.height=4}
steps_by_day_revised <- activity_revised %>%
  group_by(date) %>%
  summarize(
    total_steps = sum(steps, na.rm = TRUE)
  )

ggplot(steps_by_day_revised, aes(total_steps))+
  geom_histogram()+
  theme_minimal()+
  labs(x="Steps Taken", y="Number Days", title="Number of Days with same Step Count")

```

```{r echo=TRUE}
mean_steps_by_date_revised <- sum(activity_revised$steps,na.rm = TRUE)/length(unique(activity_revised$date))
median_steps_by_date_revised <- median(steps_by_day_revised$total_steps,na.rm=TRUE)
```

The average of steps of new data `r mean_steps_by_date_revised` is and the median is `r median_steps_by_date_revised`

## Are there differences in activity patterns between weekdays and weekends?

Creating new factor variable showing wheter data is from weekdays or weekends:
```{r echo=TRUE}
day_week <- seq_along(activity_revised$date)

for(i in 1:length(activity_revised$date)){
  if(weekdays(activity_revised$date[i]) %in% c("Saturday","Sunday")) {
    day_week[i] = "Weekend"
  } else {
    day_week[i] = "Weekday"
  }
}

day_week <- factor(day_week,c("Weekday","Weekend"))
```

Temporal distribution of steps during weekdays and weekends:
```{r echo=TRUE,fig.height=4}
activity_revised$day_week <- day_week

steps_by_intervay_and_weekday <- activity_revised %>%
  group_by(interval, day_week) %>%
  summarize(
    avg_steps = mean(steps)
  )

ggplot(steps_by_intervay_and_weekday, aes(interval,avg_steps))+
  geom_line()+
  theme_minimal()+
  facet_wrap(~day_week,nrow=2)+
  labs(x="Interval",y="Number of Steps")
```